Designing Types for R, Empirically

### 2020-05-16 paper/main.tex
Add macros
corpus

\renewcommand{\null}{\texttt{\textbf{null}}\xspace}
\newcommand{\any}{\texttt{\textbf{any}}\xspace}
\newcommand{\environment}{\texttt{\textbf{envir}}\xspace}
\newcommand{\expression}{\texttt{\textbf{expression}}\xspace}
\newcommand{\Language}{\texttt{\textbf{lang}}\xspace} %% Upper case!
\newcommand{\externalptr}{\texttt{\textbf{externalptr}}\xspace}
\renewcommand{\symbol}{\texttt{\textbf{symbol}}\xspace}
\newcommand{\pairlist}{\texttt{\textbf{pairlist}}\xspace}
\newcommand{\weakref}{\texttt{\textbf{weakref}}\xspace}
\renewcommand{\int}{\texttt{\textbf{int}}\xspace}
\newcommand{\chr}{\texttt{\textbf{chr}}\xspace}
\newcommand{\dbl}{\texttt{\textbf{dbl}}\xspace}
\newcommand{\lgl}{\texttt{\textbf{lgl}}\xspace}
\newcommand{\clx}{\texttt{\textbf{clx}}\xspace}
\newcommand{\raw}{\texttt{\textbf{raw}}\xspace}
\newcommand{\NA}{\texttt{{NA}}\xspace}
\newcommand{\VARGS}{\texttt{\textbf{...}}\xspace}
\newcommand{\T}{\ensuremath{T}\xspace}
\renewcommand{\S}{\ensuremath{S}\xspace}
\newcommand{\V}{\ensuremath{V}\xspace}
\newcommand{\A}{\ensuremath{A}\xspace}
\newcommand{\ID}{\ensuremath{ID}\xspace}
\newcommand{\F}{\ensuremath{F}\xspace}
\newcommand{\B}{\ensuremath{B}\xspace}
\newcommand{\FUN}[2]{\ensuremath{\langle #1 \rangle \rightarrow #2}\xspace}
\newcommand{\STRUCT}[1]{\texttt{\textbf{struct}}\ensuremath{\langle #1\rangle}\xspace}
\newcommand{\LIST}[1]{\texttt{\textbf{list}}\ensuremath{\langle #1\rangle}\xspace}
\newcommand{\CLASS}[1]{\texttt{\textbf{class}}\ensuremath{\langle #1\rangle}\xspace}
\newcommand{\VEC}[1]{#1\k{[]}\xspace}
\newcommand{\NAVEC}[1]{\k{\string^}\!#1\k{[]}\xspace}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\subsection{Overview of Types for R}

We ran \typetracer on the \CorpusLoadable packages we discussed in this section, which yields types for \NUMFUNCTIONS functions.
To build up a picture of what types for R functions might look like, consult Table~\ref{tbl:corpus-type-breakdown}, which lists the 10 most popular argument signatures.
The first row of the table reads: 13255 arguments have type \dbl, represnting 11.06\% of type signatures, and \typetracer observed 41M values of type \dbl flowing into or out of functions while running on the core package corpus, which represents only 1.59\% of dynamically observed values.
Of interest are rows 2 and 4, where we see that collectively \lgl and \chr make up only 16.11\% of argument types, while making up nearly 60\% of dynamically observed values.

\begin{tabular}{lrrrr}
\toprule
Type & Args with Type & \% of Static Types & Dynamic Observations & \% of Dynamic Types\\
\midrule
\dbl & 13,255 & 11.06 & 41,560,797 & 1.59\\
\lgl & 10,142 & 8.46 & 712,895,959 & 27.21\\
\null & 9,175 & 7.66 & 122,202,183 & 4.66\\
\chr & 9,171 & 7.65 & 794,058,588 & 30.31\\
\VEC\dbl & 7,811 & 6.52 & 29,272,232 & 1.12\\
\addlinespace
\VARGS & 7,759 & 6.48 & 76,369,333 & 2.92\\
\any & 6,243 & 5.21 & 255,539,497 & 9.75\\
\VEC\chr & 4,576 & 3.82 & 88,233,523 & 3.37\\
\CLASS{matrix} & 4,522 & 3.77 & 32,452,555 & 1.24\\
\CLASS{data.frame} & 2,781 & 2.32 & 7,749,094 & 0.30\\
\bottomrule
\end{tabular}

@@

\input{tables/uni-typed-breakdown.tex}
\ref{fig:singlefreq} provides a breakdown of the different type categories
observed at uni-typed positions. \scalar, \class and \vector are
the most common type categories. The shaded rows correspond to polymorphic
types. Removing those gives us XXX (XXX) monomorphic positions in a corpus of
XXX parameter and return positions.

>>>>

\renewcommand{\null}{\texttt{\textbf{null}}\xspace}
\newcommand{\scalar}{\texttt{\textbf{scalar}}\xspace}
\renewcommand{\vector}{\texttt{\textbf{vector}}\xspace}
\newcommand{\class}{\texttt{\textbf{class}}\xspace}
\newcommand{\any}{\texttt{\textbf{any}}\xspace}
\newcommand{\environment}{\texttt{\textbf{env}}\xspace}
\newcommand{\expression}{\texttt{\textbf{expression}}\xspace}
\newcommand{\Language}{\texttt{\textbf{lang}}\xspace} %% Upper case!
\newcommand{\externalptr}{\texttt{\textbf{externalptr}}\xspace}
\renewcommand{\symbol}{\texttt{\textbf{symbol}}\xspace}
\newcommand{\pairlist}{\texttt{\textbf{pairlist}}\xspace}
\newcommand{\weakref}{\texttt{\textbf{weakref}}\xspace}
\renewcommand{\int}{\texttt{\textbf{int}}\xspace}
\newcommand{\chr}{\texttt{\textbf{chr}}\xspace}
\newcommand{\dbl}{\texttt{\textbf{dbl}}\xspace}
\newcommand{\lgl}{\texttt{\textbf{lgl}}\xspace}
\newcommand{\clx}{\texttt{\textbf{clx}}\xspace}
\newcommand{\raw}{\texttt{\textbf{raw}}\xspace}
\newcommand{\NA}{\texttt{{NA}}\xspace}
\newcommand{\VARGS}{\texttt{\textbf{...}}\xspace}
\newcommand{\T}{\ensuremath{T}\xspace}
\renewcommand{\S}{\ensuremath{S}\xspace}
\newcommand{\V}{\ensuremath{V}\xspace}
\newcommand{\A}{\ensuremath{A}\xspace}
\newcommand{\ID}{\ensuremath{ID}\xspace}
\newcommand{\F}{\ensuremath{F}\xspace}
\newcommand{\B}{\ensuremath{B}\xspace}
\newcommand{\FUN}[2]{\ensuremath{\langle #1 \rangle \rightarrow #2}\xspace}
\newcommand{\STRUCT}[1]{\texttt{\textbf{struct}}\ensuremath{\langle #1\rangle}\xspace}
\newcommand{\LIST}[1]{\texttt{\textbf{list}}\ensuremath{\langle #1\rangle}\xspace}
\newcommand{\CLASS}[1]{\texttt{\textbf{class}}\ensuremath{\langle #1\rangle}\xspace}
\newcommand{\VEC}[1]{#1\k{[]}\xspace}
\newcommand{\NAVEC}[1]{\k{\string^}\!#1\k{[]}\xspace}

@@

\paragraph{Type usage}%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Table~\ref{tbl:corpus-type-breakdown} reports on the 10 most frequent types
occuring in the corpus. Scalar doubles account for 11\% function arguments,
whereas logicals represent 27\% of observed values.  Together, \lgl and \chr
account for 60\%.

\begin{table}[!h]
% BEGIN Autogenerated
\small
\begin{tabular}{lrrrr}
\toprule
Type & Args & \% of Args & Observations & \% of Observations\\
\midrule
\dbl & 13,255 & 11.06 & 41,560,797 & 1.59\\
\lgl & 10,142 & 8.46 & 712,895,959 & 27.21\\
\null & 9,175 & 7.66 & 122,202,183 & 4.66\\
\chr & 9,171 & 7.65 & 794,058,588 & 30.31\\
\VEC\dbl & 7,811 & 6.52 & 29,272,232 & 1.12\\
\addlinespace
\VARGS & 7,759 & 6.48 & 76,369,333 & 2.92\\
\any & 6,243 & 5.21 & 255,539,497 & 9.75\\
\VEC\chr & 4,576 & 3.82 & 88,233,523 & 3.37\\
\CLASS{matrix} & 4,522 & 3.77 & 32,452,555 & 1.24\\
\CLASS{data.frame} & 2,781 & 2.32 & 7,749,094 & 0.30\\
\bottomrule
\end{tabular}

@@

\input{tables/uni-typed-breakdown.tex}
\ref{fig:singlefreq} provides a breakdown of the different type categories
observed at uni-typed positions. Scalar, class and vector are
the most common type categories. The shaded rows correspond to polymorphic
types. Removing those gives us \CountMonomorphicPositions
(\PercMonomorphicPositions) monomorphic positions in a corpus of \CountPositions
parameter and return positions.
